# Reproducible Research: Peer Assessment 1

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(lubridate)
```

## Loading and preprocessing the data
##### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
unzip("activity.zip")
data <- read.csv("activity.csv")
```


        
        ## What is mean total number of steps taken per day?
        ```{r}
total_steps <- data %>%
        group_by(date) %>%
        summarise(daily_steps = sum(steps, na.rm = TRUE))
```

##### 1. Make a histogram of the total number of steps taken each day
```{r}
ggplot(total_steps, aes(daily_steps)) + geom_histogram(binwidth = 2000) +
        xlab("Total number of steps taken each day") + 
        ylab("Frequency")
```

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
mean <- mean(total_steps$daily_steps, na.rm=TRUE)
median <- median(total_steps$daily_steps, na.rm=TRUE)
```
* Mean: `r mean`
* Median:  `r median`

-----
        
        ## What is the average daily activity pattern?
        ```{r}
interval_steps <- data %>% 
        group_by(interval) %>%
        summarise(steps = mean(steps, na.rm =TRUE))

```

##### 1. Make a time series plot
```{r}
ggplot(data=interval_steps, aes(x=interval, y=steps)) +
        geom_line() +
        xlab("5-minute intervals") +
        ylab("Average number of steps taken")
```

        ## Imputing missing values
        ##### 1. Calculate and report the total number of missing values in the dataset 
        ```{r}
missing <- !complete.cases(data)
```

* Number of missing values: `r missing`

## impute missing steps with interval averages across days
```{r}
imputed_data <- data %>%
        mutate(
                steps = case_when(
                        is.na(steps) ~ interval_steps$steps[match(data$interval, interval_steps$interval)],      
                        TRUE ~ as.numeric(steps)
                ))
```

##### 4. Make a histogram of the total number of steps taken each day 
```{r}

imputed_total_steps <- imputed_data %>% group_by(date) %>% summarise(daily_steps = sum(steps))

ggplot(imputed_total_steps, aes(daily_steps)) + 
        geom_histogram(binwidth = 2000) + 
        xlab("Total number of steps taken each day") + 
        ylab("Frequency")


```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
imputed_mean = mean(imputed_total_steps$daily_steps, na.rm=TRUE)
imputed_median = median(imputed_total_steps$daily_steps, na.rm=TRUE)

mean_diff <- imputed_mean - mean 
median_diff <- imputed_median - median

```
----
        
        ## Are there differences in activity patterns between weekdays and weekends?
        ##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
        
        ```{r}
day_of_week <- imputed_data %>%
        mutate(
                date = ymd(date),
                weekday_or_weekend = case_when(wday(date) %in% 2:6 ~ "Weekday",
                                               wday(date) %in% c(1,7) ~ "Weekend")
        ) %>% select(-date) %>%
        group_by(interval, weekday_or_weekend) %>%
        summarise(
                steps = mean(steps)
        )
```

##### 2. Make a panel plot containing a time series plot

```{r}

ggplot(day_of_week, aes(interval, steps)) + 
        geom_line() + 
        facet_wrap(~weekday_or_weekend, nrow = 2) +
        xlab("5-Minute intervals") + 
        ylab("Average number of steps")
```
